### 2024-04-18 project3/tlHedging.tex
new

\chapter{Efficient Transfer of Knowledge for Deep Hedging of Variable Annuities}

@@

In this paper, we propose a transfer learning approach to improve the risk management of variable annuities (VAs) by learning a hedging strategy utilizing previous experiences from other financial products.
VAs are popular retirement products that combine investment and insurance features, offering policyholders the potential for investment growth and the protection of a guaranteed minimum death benefit or income benefit.
Due to the complexity of the products and the dynamic nature of the financial markets, the task of hedging VAs is particularly challenging.
We aim to leverage the knowledge from the hedging strategies of financial derivatives to construct hedging strategies for VAs, thereby improving the risk management of these products.
The rest of the paper is organized as follows: Section~\ref{sec3:vaHedging} presents the problem formulation for hedging variable annuity with a deep neural network, Section~\ref{sec3:tlHedging} presents the background knowledge for transfer learning, and Section~\ref{sec3:numerical} states the plans for ongoing numerical experiments.

@@

\subsection{Markov Decision Process (MDP) for Hedging VAs}
A Markov Decision Process (MDP) provides a mathematical framework for solving sequential decision-making tasks in situations where outcomes are partly random and partly under the control of an agent. 
In the context of hedging VAs, an MDP can be used to model the decision-making process of an insurer who aims to minimize the liability of the VA by dynamically adjusting the hedging portfolio based on the current state of the VA and the financial markets.
An MDP is defined by the tuple $\mathcal{M} = (\mu_0, \mathcal{S}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$, where:

where $R_{t+1} = \mathcal{R}(s_i, a_i, s_{i+1})$ is the reward received by the agent after taking action $a_i$ in state $s_i$ and transitioning to state $s_{i+1}$, and the expectation is taken over the possible sequences of states and rewards that follow from the policy $\pi$.
In the context of hedging VAs, the initial states $\mu_0$ specifies the initial value of the VA and the market specifications. 
The state space $\mathcal{S}$ represent the information available to the agent, such as current value of the underlying asset ($S_t$), the current subaccount value ($F_t$), and the guarantee value ($G_t$), as well as other relevant contract specifications.
The action space $\mathcal{A}$ represent the hedging decisions, such as the hedging portfolio weights.
The state transition probability distribution $\mathcal{P}$ and the reward function $\mathcal{R}$ are derived from the dynamics of the VA and the market specifications. Instead of having an infinite time horizon, the final reward is obtained at the maturity of the VA contract. Hence, $R_{t} = 0$ for all $t > T$.
The discount factor $\gamma$ models the present value of future rewards and is conveniently inherited from the risk-free rate.

@@

Depending on the specific market model and the hedging objective, deep hedging can be categorized into model-based and model-free approaches.
Model-based deep hedging trains a deep neural network (DNN) to learn optimal hedging actions by maximizing a value function that reflects the hedging error or risk measure, such as the variance of the hedging portfolio's final value or tail risk measures like the CVaR.
The DNN processes sequential market data and outputs a policy that determines the hedging actions based on the current state of the market.
In the context of deep hedging, the policy $\pi = \phi(\mathcal{I})$ is defined by the parameter $\phi$ of a DNN, and the information $\mathcal{I}$ is the training data during the learning process.
~\cite{buehler2019deep} is the first paper to propose a model-based deep hedging approach for hedging financial derivatives with a focus on minimizing CVaR.
A more recent paper by~\cite{imaki2021no} proposes a new network architecture for model-based deep hedging that can be trained more efficiently.
A typical setup for a model-based deep hedging strategy includes:

@@

@@

    A policy gradient method uses policy gradient theorem to update the policy by performing gradient ascent with respect to the policy parameters.
    Representative algorithms include the Proximal Policy Optimization (PPO) algorithm proposed by~\cite{schulman2017proximal} and the Trust Region Policy Optimization (TRPO) algorithm proposed by~\cite{schulman2015trust}.
    The most relevant paper to our study is~\cite{chong2023pseudo}, which uses the PPO algorithm to hedge GMMB with a GMDB rider under the Black-Scholes framework.
\end{enumerate}

\section{Transfer Learning for Risk Management of Variable Annuities} \label{sec3:tlHedging}

Transfer learning is a machine learning technique that leverages knowledge from one domain to improve learning in another domain.
In the context of deep reinforcement learning, transfer learning can be used to improve the risk management of VAs in five ways:
\begin{enumerate}
    \item   \textbf{Transferring to VA products}: Knowledge can be transferred from the hedging strategies of financial derivatives to constructing hedging strategies for VAs.
    VAs are complex insurance products with unique features, which makes it challenging to develop effective hedging strategies.
    However, the payoff structures of VAs are similar to those of financial derivatives, such as European options, which have been extensively studied in the literature.
    By leveraging the knowledge from the hedging strategies of financial derivatives, we can improve the risk management of VAs.
    \item   \textbf{Transferring to new VA products}: Knowledge can be transferred from the hedging strategies of one VA product to another VA product.
    Different VA products have different features and payoff structures, which makes it challenging to develop hedging strategies for new VA products.
    However, experience from hedging one VA product can be transferred to develop hedging strategies for new VA products, especially when the new VA product is similar to the existing ones.
    \item   \textbf{Transferring to new asset models}: Knowledge can be transferred from the hedging strategies of one asset model to another asset model.
    A stochastic volatility model is a more realistic model for asset prices than the Black-Scholes model, but it is more challenging to develop hedging strategies for VAs under such a model.
    \item  \textbf{Transferring to new model specifications}: Model misspecification is a common issue in financial modeling, and training on data simulated with misspecified parameters can lead to suboptimal hedging strategies.
    If only the misspecified model is similar to the correct model, then we can leverage the knowledge from the misspecified model to improve the hedging strategies under the correct model, since the only difference between the two models is the parameter values. 
    \item   \textbf{Transferring to the real world}: Knowledge can be transferred from simulated data to real-world data.
    In practice, it is challenging to obtain real-world data for VAs, but we can leverage simulated data to develop hedging strategies for VAs.
    If the simulated data is representative of the real-world data, then the hedging strategies developed from the simulated data can be transferred to the real world.

@@

Given a set of source domains $\bm{\mathcal{M}}_s= \{ \mathcal{M}_s| \mathcal{M}_s \in \bm{\mathcal{M}}_s\}$ and a target domain $\mathcal{M}_t$, transfer learning aims to improve the learning on the target domain.
While utilizing the interior information $\mathcal{I}_t$ from the target domain, the exterior information $\mathcal{I}_s$ from the source domains is also leveraged to improve the learning on the target domain.

\begin{equation}
    \pi^* = \arg \max_\pi \mathbb{E}_{s \sim \mu_0, a \sim \pi} \left[V_{\mathcal{M}}^\pi (s, a) \right], 
\end{equation}

where $\pi = \phi(\mathcal{I}_s \sim \bm{\mathcal{M}}_s, \mathcal{I}_t \sim \mathcal{M}_t)$ is policy learned from target domain based on information from both the source domains and the target domain.
A regular reinforcement without transfer learning is a special case where $\mathcal{I}_s = \emptyset$ and $\pi = \phi(\mathcal{I}_t \sim \mathcal{M}_t)$.

In the context of deep hedging, various research has been conducted to improve the hedging strategies on real-world data by leveraging knowledge from simulated data.
Most of the research uses the offline-online learning approach, where the hedging strategies are first trained with simulation data, but updated in real-time based on new market information and continuous learning to adapt to changing market dynamics.
The most relevant paper to our study is~\cite{xiao2021optimal}, which use policy gradient algorithms to train hedging strategies for European options with simulation data from Heston model and then updated in real-time hedging S\&P 500 index options.
To the best of our knowledge, no research has been conducted to improve the hedging strategies with the state-of-the-art transfer learning techniques.
This section introduces some transfer learning techniques that are relevant to our task of hedging VAs.

\subsection{Reward Shaping}

In the context of hedging VAs using deep reinforcement learning and transfer learning, reward shaping becomes a crucial technique. 
It involves modifying the reward function in the target domain to incorporate additional guidance, which often leads to more efficient learning and better performance.
This technique can be particularly useful for managing the complex risks associated with VAs, as it can guide the learning process towards strategies that are more effective in hedging these products.

One way to implement reward shaping in this context is by introducing auxiliary rewards that reflect the performance of hedging strategies learned from the source domains.
Reward shaping learns an auxiliary reward function $\mathcal{R}_s: \mathcal{S} \times \mathcal{A} \times \mathcal{S} \rightarrow \mathbb{R}$ from the source domains.
The auxiliary reward function is then used to shape the reward function in the target domain, $\mathcal{R}'_t = \mathcal{R}_t + \mathcal{R}_s$, by incorporating the insights gained from the source domains.

Some popular reward shaping techniques include potential-based reward shaping~\citep{ng1999policy}, potential based state-action advice~\citep{wiewiora2003principled}, dynamic value function advice~\citep{harutyunyan2015expressing}.

\subsection{Policy Transfer}

The most straightforward approach to transfer learning is to directly transfer the policy learned from the source domains to the target domain via policy distillation.
Policy distillation is a technique that transfers the knowledge from a teacher policy to a student policy by training the student policy to mimic the teacher policy up to a certain extent.
The teacher policy is learned from the source domains, and the student policy learns the hedging strategies in the target domain while guided by the teacher policy.
In the Distral algorithm~\citep{teh2017distral}, the student policy maximizes a multi-task objective:
$\max_{\phi} \sum_{i=1}^K \mathcal{J}(\pi_\phi, \pi_{E_i})$, where

\begin{equation}
    \mathcal{J}(\pi_\phi, \pi_{E_i}) = \sum_{t=0}^\infty \mathbb{E}_{s_t \sim \mu_0^t, a_t \sim \pi_\phi} \left[ \gamma^t R_{t+1} + \frac{\alpha}{\beta} \log \pi_{\phi}(a_t|s_t) - \frac{1}{\beta}log \pi_{E_i}(a_t|s_t) \right],
\end{equation}

where a set of $K$ teacher policies $\pi_{E_i}$ are learned from the source domains, and the student policy $\pi_\phi$ is learned on the target domain. $\alpha$ and $\beta$ are hyperparameters that control the trade-off between mimicking the teacher policies and exploring the target domain.

\subsection{Representation Transfer}

Representation transfer is a technique that transfers the knowledge based on the following critical assumption:

\begin{assumption}
    The state space $\mathcal{S}$, the action space $\mathcal{A}$, and the reward function $\mathcal{R}$ can be disentangled into orthogonal subspaces, which are task-invariant such that knowledge can be transferred between domains on the universal subspaces.
\end{assumption}

A suitable representation transfer technique for our task of hedging VAs is modular networks~\citep{andreas2017modular}, which decomposes a policy $\pi$ into two sub-modules $g$ and $f$, i.e.,

\begin{equation}
    \pi = \phi(s_{market}, s_{guarantee}) = f(g(s_{market}), s_{guarantee}),
\end{equation}

~\cite{zhang2018decoupling} uses modular networks to transfer knowledge of physics models from previous tasks to new tasks in a simulated environment.
In the context of hedging VAs, $g$ is the module that processes the market information, and $f$ is the module that processes the guarantee information.
The core idea behind modular networks is that the market information can be applied for hedging strategies across different VA products, while the guarantee information is specific to each VA product.
Hence, knowledge can be transferred from the source domains to the target domain by transferring the market module $g$.

\subsection{Evaluation Metrics}

The metrics discussed for evaluating transfer learning approaches focus on two key aspects: mastery and generalization. 
Mastery assesses the agent's final performance level in the target domain, indicating how effectively it has learned a specific task from previous knowledge. 
Some common metrics for evaluating mastery include:

\begin{itemize}
    \item \textbf{Asymptotic performance:} the final reward of the agent in the target domain.
    \item \textbf{Transfer ratio:} the ratio of the agent's performance in the target domain with and without transfer learning.
\end{itemize}

Generalization, on the other hand, refers to the agent's ability to quickly adapt its learned knowledge to the target domain.

\begin{itemize}
    \item \textbf{Jumpstart performance:} the initial reward of the agent in the target domain.
    \item \textbf{Accumulated reward:} the total area under the reward curve over time in the target domain.
    \item \textbf{Time to threshold:} the time it takes for the agent to reach a certain performance threshold in the target domain.
    \item \textbf{Performance with fixed epochs:} the agent's performance in the target domain after a fixed number of epochs.
\end{itemize}

\section{Plans for Ongoing Numerical Experiments} \label{sec3:numerical}

As we aim to enhance the risk management strategies for variable annuities (VAs) through the application of transfer learning, our next step involves conducting a series of numerical experiments. These experiments are crucial for validating the theoretical frameworks and methodologies developed so far. Our plan includes:

\begin{enumerate}
    \item \textbf{Simulation setup:} Designing a simulation environment that closely mirrors real-world market dynamics and VA contract specifics. 
    This setup will include various asset models, including the Black-Scholes and stochastic volatility models, to test the adaptability of transferred hedging strategies across different market conditions.
    \item \textbf{Baseline model comparison:} Establishing a set of baseline hedging strategies against which we can compare the performance of our transfer learning-enhanced strategies. 
    This will likely involve traditional delta hedging techniques and possibly other reinforcement learning approches that have not utilized transfer learning.
    \item \textbf{Transfer learning implementation:} Applying the transfer learning methodologies discussed, including reward shaping, policy transfer, and/or representation transfer to the task of hedging VAs. 
    The transfer can happen between different VA products, different asset models, different model specifications, and between simulated and real-world data.
    \item \textbf{Performance evaluation:} Utilizing the evaluation metrics outlined in the previous sections to assess mastery and generalization of the transfer learning strategies. 
    Key metrics will include asymptotic performance, transfer ratio, jumpstart performance, and time to threshold.
    \item \textbf{Sensitivity analysis:} Conducting sensitivity analyses to understand how changes in market conditions, contract specifics, and model parameters affect the performance of our transfer learning strategies. This will help in identifying the robustness and limitations of our approach.
    \item \textbf{Hedging in real-world:} If accessible, applying the trained models to real-world data to further validate their practical applicability and performance in real-world market conditions.
\end{enumerate}

Through these numerical experiments, we aim to demonstrate the efficacy of transfer learning in improving the hedging of VAs in a variety of market conditions and contract specifics. 
We expect that our transfer learning-enhanced strategies will outperform traditional hedging techniques and possibly other machine learning approaches that have not utilized transfer learning. 
This will provide strong evidence for the potential of transfer learning in improving risk management strategies for VAs and other complex financial products in a dynamic market environment.

\subsection{Initial Numerical Experiments}

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\textwidth]{./project3/figures/Figure_1.png}
    \caption{Initial numerical experiments for GMMB}
    \label{fig3:GMMB}
\end{figure}

>>>>

\chapter{Model-Free Deep Hedging of Variable Annuities}

@@

In this paper, we propose a model-free RL algorithm to improve the risk management of variable annuities (VAs) by learning a hedging strategy from past experience.
VAs are popular retirement products that combine investment and insurance features, offering policyholders the potential for investment growth and the protection of a guaranteed minimum death benefit or income benefit.
Due to the complexity of the products and the dynamic nature of the financial markets, the task of hedging VAs is particularly challenging.
Our model-free RL approach aims to learn an optimal hedging policy for VAs without relying on a predefined model of the underlying asset dynamics.
By leveraging the flexibility and adaptability of RL algorithms, we aim to develop a dynamic hedging strategy that can effectively manage the risks associated with VAs in a changing market environment.

The rest of the paper is organized as follows: Section~\ref{sec3:vaHedging} presents the problem formulation for hedging variable annuity with a deep neural network,

@@

@@

Depending on the specific market model and the hedging objective, deep hedging can be categorized into model-based and model-free approaches.
Model-based deep hedging trains a deep neural network (DNN) to learn optimal hedging actions by maximizing a value function that reflects the hedging error or risk measure, such as the variance of the hedging portfolio's final value or tail risk measures like the CVaR.
The DNN processes sequential market data and outputs a policy that determines the hedging actions based on the current state of the market.
In the context of deep hedging, the policy $\pi = \phi(\mathcal{I})$ is defined by the parameter $\phi$ of a DNN, and the information $\mathcal{I}$ is the training data during the learning process.
~\cite{buehler2019deep} is the first paper to propose a model-based deep hedging approach for hedging financial derivatives with a focus on minimizing CVaR.
~\cite{carbonneau2021deep} extends the model-based approach to hedging long-term financial derivatives and compares the performance of deep hedging strategies trained with different reward functions.
A more recent paper by~\cite{imaki2021no} proposes a new network architecture for model-based deep hedging that can be trained more efficiently by implementing a no-tranaction band.
A typical setup for a model-based deep hedging strategy includes:

@@

In summary, model-based deep hedging approaches rely on explicit market models to learn the optimal hedging strategies.
It requires the knowing of the underlying asset price dynamics and the risk factors affecting the financial instruments, which may limit its flexibility and robustness to model misspecification.
~\cite{chong2023pseudo} specifically discuss the effect of model misspecification on the performance of model-based deep hedging strategies for hedging variable annuities with a guaranteed minimum death benefit (GMDB) rider under the Black-Scholes framework.

@@

    A policy gradient method uses policy gradient theorem to update the policy by performing gradient ascent with respect to the policy parameters.
    Representative algorithms include the Trust Region Policy Optimization (TRPO) algorithm proposed by~\cite{schulman2015trust} and the Proximal Policy Optimization (PPO) algorithm proposed by~\cite{schulman2017proximal}.

    The most relevant paper to our study is~\cite{chong2023pseudo}, which uses the PPO algorithm to hedge GMMB with a GMDB rider under the Black-Scholes framework.
    However,~\cite{chong2023pseudo} implement the PPO algorithm with the knowing of the underlying asset model and the risk factors affecting the mortality rate, while our study focuses on the model-free deep reinforcement learning approach to hedge VAs without explicit model assumptions.

@@

\section{Proximal Policy Optimization for Hedging Variable Annuities}

PPO is a policy gradient method that aims to optimize the policy by performing gradient ascent with respect to the policy parameters.
Originally proposed by~\cite{schulman2017proximal} and inspired by TRPO, PPO aims to address the limitation of a crude policy gradient method by introducing a clipped surrogate objective function that prevents large policy updates and ensures more stable training.
While TRPO uses a hard coded constraint on the policy update to ensure that the new policy does not deviate too far from the old policy, PPO uses a clipped surrogate objective function that penalizes a high Kullback–Leibler (KL) divergence between the new and old policies. 
This allows for more significant policy updates while maintaining the stability of the training process.

The PPO algorithm consists of two main components: the policy network and the value network.
The policy network learns the policy by outputting the action based on the current state, while the value network learns the value function by approximating the expected cumulative reward over time following the policy.
The policy network is updated by performing gradient ascent with respect to the policy parameters, while the value network is updated by minimizing the mean squared error between the predicted value and the target value.

The PPO algorithm is particularly well-suited for continuous action spaces and has been successfully applied to a wide range of reinforcement learning tasks, including robotic control, game playing, and financial trading.
In the context of hedging VAs, PPO can be used to learn an optimal hedging policy by maximizing a reward function that reflects the hedging performance.

\subsection{Markov Decision Process (MDP) for Hedging VAs}
A Markov Decision Process (MDP) provides a mathematical framework for solving sequential decision-making tasks in situations where outcomes are partly random and partly under the control of an agent. 
In the context of hedging VAs, an MDP can be used to model the decision-making process of an insurer who aims to minimize the liability of the VA by dynamically adjusting the hedging portfolio based on the current state of the VA and the financial markets.
In this section, we reformulate the hedging problem of VAs from Section~\ref{subsec:VApayout} as an MDP and define the key components of the MDP.
An MDP is defined by the tuple $\mathcal{M} = (\mu_0, \mathcal{S}, \mathcal{A}, \mathcal{P}, \mathcal{R}, \gamma)$, where:

where $R_{t+1} = \mathcal{R}(s_i, a_i, s_{i+1})$ is the reward received by the agent after taking action $a_i$ in state $s_i$ and transitioning to state $s_{i+1}$, and the expectation is taken over the possible sequences of states and rewards that follow from the policy $\pi$.

In the context of hedging VAs, the initial states $\mu_0$ specifies the initial value of the VA and the market specifications. 
The state space $\mathcal{S}$ represent the information available to the agent, such as current value of the underlying asset ($S_t$), the current subaccount value ($F_t$), and the guarantee value ($G_t$), as well as other relevant contract specifications.
The action space $\mathcal{A}$ represent the hedging decisions, such as the hedging portfolio weights.
The state transition probability distribution $\mathcal{P}$ and the reward function $\mathcal{R}$ are derived from the dynamics of the VA and the market specifications. Instead of having an infinite time horizon, the final reward is obtained at the maturity of the VA contract. Hence, $R_{t} = 0$ for all $t > T$.
The discount factor $\gamma$ models the present value of future rewards and is conveniently inherited from the risk-free rate.
